---
title: "A3"
output: pdf_document
---

```{r setup, include=FALSE}
library(dplyr)
library(knitr)
library(tidyverse)
library(plm)
library(fastDummies)
library(sandwich)

knitr::opts_chunk$set(echo = FALSE)
```

## EX1 Data Set
Load dataset.
```{r loaddata, echo = FALSE}
pop = read.csv("population.csv")
crime = read.csv("crime_long.csv")
officers = read.csv("officers.csv")
```

## EX2 Data Manipulation

$\bullet$ Calculate total crime per month and plot the time series of crime.

```{r  echo=FALSE}
crime_per_month = crime %>% select(crime_month, crimes)%>% group_by(crime_month) %>% summarise(Crime = sum(crimes))
crime_per_month$crime_month = as.Date(crime_per_month$crime_month, "%Y-%m-%d")
plot(crime_per_month$Crime ~ crime_per_month$crime_month, crime_per_month, xaxt = "n", type = "l",
     xlab = "Month", ylab = "Number of Crimes", main = "Total Crime per Month")
axis(1, crime_per_month$crime_month,format(crime_per_month$crime_month, "%Y/%m"))

```

$\bullet$ Merge two datasets by district-units and period.
```{r echo = FALSE}
crime= crime %>% group_by(crime_month, district, crime_type) %>% summarise(crimes = sum(crimes), .groups ='keep')
data = left_join(crime, pop, by = c("crime_month" = "month", "district" = "district"))
# remove rows with NA
data = data[complete.cases(data), ]
head(data)
```
$\bullet$ Construct a panel data of unit over time.
```{r echo=FALSE}
# districts over time
panel = unique(data %>% ungroup()%>% select(crime_month, period, district))
# total crimes per resident
crimes_per_pl = data %>% ungroup() %>% select(crime_month, district, crimes, tot_pop)%>% group_by(crime_month, district) %>% summarise(crimes_per_pl = sum(crimes)/mean(tot_pop), .groups = 'keep')
panel = panel%>% left_join(crimes_per_pl, by = c("crime_month", "district"))
# violent crimes per resident
viocrimes_per_pl = data %>% filter(crime_type == "violent") %>% ungroup() %>% select(crime_month, district, crimes, tot_pop)%>% group_by(crime_month, district) %>% summarise(viocrimes_per_pl = sum(crimes)/mean(tot_pop), .groups = 'keep')
panel = panel%>% left_join(viocrimes_per_pl, by = c("crime_month", "district"))
# property crimes per resident
procrimes_per_pl = data %>% filter(crime_type == "property") %>% ungroup() %>% select(crime_month, district, crimes, tot_pop)%>% group_by(crime_month, district) %>% summarise(procrimes_per_pl = sum(crimes)/mean(tot_pop), .groups = 'keep')
panel = panel%>% left_join(procrimes_per_pl, by = c("crime_month", "district"))
# median income
medincome = data %>% ungroup() %>% select(crime_month, district, p50_inc) %>% 
  group_by(crime_month, district) %>% summarise(medinc = mean(p50_inc), .groups = 'keep')
panel = panel%>% left_join(medincome, by = c("crime_month", "district"))
```
```{r echo=FALSE}
# share of black, Hispanic, and white residents
shares = data %>% ungroup() %>% select(crime_month, district, tot_pop, tot_white, tot_black, tot_hisp) %>% group_by(crime_month, district) %>% summarise(perc_white = mean(tot_white)/mean(tot_pop), perc_black = mean(tot_black)/mean(tot_pop), perc_hisp = mean(tot_hisp)/mean(tot_pop), .groups = 'keep')
panel = panel %>% left_join(shares, by = c("crime_month", "district"))
```
```{r echo=FALSE}
head(panel)
```
```{r echo =  FALSE}
# remove variables no longer needed
rm(crime_per_month, crimes_per_pl, medincome, shares, procrimes_per_pl, viocrimes_per_pl, pop)
```


## EX3 Panel Data: Introduction
```{r echo = FALSE}
# create dataset
data1 = officers %>% left_join(panel, by = c("month" = "crime_month", "unit" = "district"))
```
$\bullet$ With intercept.
```{r echo = FALSE}
# regression (with intercept)
model1 = lm(arrest ~ tenure + crimes_per_pl + medinc + perc_black + perc_hisp + perc_white, data = data1)
summary(model1)
```
$\bullet$ Without intercept.
```{r echo=FALSE}
# Without intercept
model2 = lm(arrest ~ tenure + crimes_per_pl + medinc + perc_black + perc_hisp + perc_white - 1, data = data1)
summary(model2)
```

## EX4 Panel Data: More controls
$\bullet$ Creating dummy by hand.
```{r echo=FALSE}
data2 = data1
# generate dummy variables for districts
data2 = fastDummies::dummy_cols(data2, select_columns = "unit") %>% select(-unit)
# generate dummy variables for year and month
data2 = fastDummies::dummy_cols(data2, select_columns = "month") %>% select(-month)
# remove variables we don't use
data2 = data2 %>% select(-c(NUID, period, viocrimes_per_pl, procrimes_per_pl))
data2 = data2[,-33] # drop the first year-month: 2007-01-01
# regression with dummies
model3 = lm(arrest ~. - 1, data = data2)
summary(model3)
```

$\bullet$ Using factor() yields the same result.
```{r echo=FALSE}
# using the lm model with factor() for fixed effects
model4 = lm(arrest ~ tenure+ crimes_per_pl+medinc+perc_black+perc_hisp+perc_white+factor(unit)+factor(month) - 1, data = data1)
summary(model4)
```

## EX5 Panel Data: Individual fixed effects
Encounter memory error when creating dummy for individual fixed effects, so I use equation (2).
$\bullet$ Within estimator for $\beta$. Since in EX4 we can see the results by using factor() and generating dummies by hand are the same, I used factor() in lm for fixed effects in this question. 
Encounter memory error when creating dummy for individual fixed effects.
```{r echo=FALSE}
model5 = lm(arrest ~ tenure+ crimes_per_pl+medinc+perc_black+perc_hisp+perc_white +factor(unit)+factor(month) - 1, data = data1)
summary(model5)
```
$\bullet$ Between estimator of $\beta$.
```{r echo=FALSE}
# group data to average out the time component
data_avgtime = data1 %>% group_by(NUID)%>% summarise(unit = mean(unit), tenure = mean(tenure), arrest = mean(arrest), totcrime = mean(crimes_per_pl), medin = mean(medinc), black = mean(perc_black), hisp = mean(perc_hisp), white = mean(perc_white))
# run estimation
model6 = lm(arrest ~ tenure+ totcrime +medin + black+ hisp+ white + factor(unit) - 1, data = data_avgtime)
summary(model6)
```

$\bullet$ First difference estimator for $\beta$.
```{r echo=FALSE}
data_sort = data1[order(data1$NUID, data1$month),]
n = unique(data1$NUID)
n = n[!is.na(n)]
beta = mat.or.vec(1, length(n))
j = 0
#for (i in n){
#  dat = data_sort%>%filter(NUID == i)
#  if (nrow(dat)>1){
#    df = data.frame(diff(as.matrix(dat%>%select(-month))))
#  }else{
#    df = dat
#  }
#  m = lm(arrest~tenure+ crimes_per_pl+medinc+perc_black+perc_hisp+perc_white -1, data = df)
#  beta[j] = m$coefficients[["tenure"]]
#  j = j + 1
#}
# mean(beta)
```
```{r echo=FALSE}
# GMM

```





